Shapes analysis for time series.

### 2024-05-15 NEURIPS24/neurips_2024_to_submit.tex
update

\begin{abstract}
    %TO CHANGE]
    Analyzing inter-individual variability of physiological functions is particularly appealing in medicine and biology to describe or quantify health conditions.
     Such analysis can be done by comparing individuals to a reference one using their time series as biomedical data.
     % In biomedical research, time series data is a prevalent source of information, necessitating dedicated methods comparing them and averaging them to derive a valuable reference time series. 
 This paper introduces an unsupervised representation learning (URL) algorithm for time series tailored to biomedical inter-individual studies.
 The idea is to represent time series as deformations of a reference time series. The deformations are diffeomorphisms parameterized and learned by our method called TS-LDDMM.
  Once the deformations and the reference time series are learned, the vector representations of individual time series are given by the parametrization of their corresponding deformation.
   At the crossroads between URL for time series and shape analysis, the proposed algorithm handles irregularly sampled multivariate time series of variable lengths and provides shape-based representations of temporal data. 
 In this work, we establish a representation theorem for the graph of a time series and derive its consequences on the LDDMM framework. We showcase the advantages of our representation compared to existing methods using synthetic data and real-world examples motivated by biomedical applications.
\end{abstract}

>>>>

\begin{abstract}
    %TO CHANGE]
    Analyzing inter-individual variability of physiological functions is particularly appealing in medical and biological contexts to describe or quantify health conditions. Such analysis can be done by comparing individuals to a reference one with time series as biomedical data. 
This paper introduces an unsupervised representation learning (URL) algorithm for time series tailored to inter-individual studies. The idea is to represent time series as deformations of a reference time series. The deformations are diffeomorphisms parameterized and learned by our method called TS-LDDMM. Once the deformations and the reference time series are learned, the vector representations of individual time series are given by the parametrization of their corresponding deformation. At the crossroads between URL for time series and shape analysis, the proposed algorithm handles irregularly sampled multivariate time series of variable lengths and provides shape-based representations of temporal data.
 In this work, we establish a representation theorem for the graph of a time series and derive its consequences on the LDDMM framework. We showcase the advantages of our representation compared to existing methods using synthetic data and real-world examples motivated by biomedical applications.
\end{abstract}

### 2024-05-15 NEURIPS24/paper.tex
update

 A time series graph can be seen a curve and LDDMM structure is relevant to tackle curves as presented in \cite{glaunes2008large}. %, LDDMM structure is relevant to tackle curves and
  However, time series graph has more structure as depicted in \Cref{fig:diffeo} due to the temporal evolution. 
  \cite{qiu2009time} tracks anatomical shape changes in serial images using LDDMM, but distinguish from us by including the temporal evolution at a higher level: the goal is to perform longitudinal data modeling.

@@

   Very recently in a functionnal data analysis framework \cite{wu2024shape} (Shape-FPCA) improved by giving also a representation for the original time evolution.

>>>>

 A time series graph can be seen as a curve and LDDMM structure is relevant to tackle curves as presented in \cite{glaunes2008large}. %, LDDMM structure is relevant to tackle curves and
  However, time series graph has more structure than curves as depicted in \Cref{fig:diffeo} due to the temporal evolution. 
  \cite{qiu2009time} tracks anatomical shape changes in serial images using LDDMM, but distinguish from us by including the temporal evolution at a higher level: the goal is to perform longitudinal data modeling.

@@

   Very recently in a functionnal data analysis framework, a paper \cite{wu2024shape} (Shape-FPCA) improved by giving a representation for the original time evolution.
